Limits of Media Effects

### 2022-09-29 paper/Manuscript_Framing.Rmd
Update: Limits of Media Effects

---
title: "Framing Effects in the Wild"
author: "Nicolai Berk"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    fig_caption: yes
    # includes:
    #   in_header: preamble.tex
# subtitle: Manuscript prepared for the 12th Annual Conference of the European Political Science Association in Prague, Czech Republic
fontsize: 12pt
geometry: margin=1.5in
abstract: "Does news framing affect public opinion? While a large body of experimental work suggests that individual attitudes can be easily altered by emphasizing different aspects of an issue, little is known about how these studies generalize to real world environments. I exploit a rare change in the migration framing of the major German tabloid *Bild* to bridge this gap. In a first step, I develop a framework for the measurement of emphasis frames in political communication using supervised machine learning. Using this framework, I show that the emphasis of crime in *Bild*'s migration coverage severely increased following an exogenously timed change of editor. Second, I employ a difference-in-differences design using panel data from over 16,000 individuals in thirteen waves to assess the precise impact of this shift on migration attitudes. Despite a major change in migration framing, a robust null effect on immigration attitudes and a number of related variables can be observed. These findings caution against the generalization of experimental framing studies to real-world environments and open a broader research agenda to study the precise impact of political communication beyond the experimental setting."
bibliography: C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib
nocite: |
  @CologneReport2016
indent: true
---

```{r setup, include = F}

knitr::opts_chunk$set(echo = F, message = F, warning = F, cache = T, fig.height=6, fig.width=10, fig.align='center', eval = T, fig.pos = "!ht", out.extra = "")

library(tidyverse)
library(dplyr)
library(data.table)
library(fixest)
library(lubridate)
library(patchwork)
library(here)

@@

*I am thankful for extensive feedback and other support on this manuscript from Christian Arnold, Daniel Bischof, Philipp Broniecki, Jóhanna Ýr Bjarnadóttir, Elias Dinas, Florian Foos, Violetta Haas, Felix Hagemann, Anselm Hager, Julian Heid, Christina Gahn, Heike Klüver, Martin Vinæs Larsen, Matthew Levendusky, Will Lowe, Thomas M. Meyer, Rune Slothuus, Judith Spirig, Antonio Valentim, as well as participants of CompText 2022, EPSA 2022, ICA 2022, EPSIP Colloquium at HU Berlin, CIVICA EGN Workshop at the EUI,  and the Political Behaviour Section at Aarhus University. I also thank the two research assistants, Robin Groß and Nathalie Aerts, for annotating the news articles. All remaining confusions are my own.*

# Introduction: Does Framing Matter?

Does news framing affect public opinion? Despite the profound importance for the understanding of elite influence in modern democracies, we know little about the persuasive effects of news framing in real-world environments. A large body of experimental evidence suggests that the emphasis of specific considerations can substantially alter issue attitudes, However observational data suggests little variation in individual attitudes [@Kustov2021] and prior work has shown that the external validity of such experiments might be seriously complicated by the current salience and politicization of the issue of interest [@Arceneaux2009; @Bechtel2015; @Lecheler2009b], the availability of different frames [@Chong2007a; @Chong2013; @Druckman2003; @Druckman2012] and communicating actors [@Slothuus2010], as well as the seeking and avoidance of specific content [@Arceneaux2009; @Lau2021a; @Taber2006]. Contemporary news environments provide consumers with especially diverse content and maximize their abilities to choose specific outlets and avoid others [@Bennett2008], thereby increasing the gap between the real-world environment and experimental studies. Additionally, the very nature of experiments is prone to over-estimate effects in the real world [@Barabas2010]. This is the reason why, fifteen years ago, Donald Kinder called for framing research to extend its toolkit: "[E]xperimental results can always be questioned on their generalizability, and framing effects are no exception. [...] A more balanced reading of frame effects requires methodological diversification, experiments and studies oriented to the world outside" [-@Kinder2007, 157].

In this article, I provide the - to my knowledge - first observational estimate of framing effects by exploiting a rare shift in the framing of a highly politicized issue - migration - in Germany's largest tabloid newspaper, *Bild*. The case is ideal to assess the everyday influence of issue frames on public opinion, as one outlet severely changed its coverage, while other papers did not. Providing a novel framework for the measurement of emphasis frames using BERT models, I study migration framing in a corpus of 2.5 million newspaper articles. I show that the newspaper's migration coverage saw a strong increase in the use of crime frames after the editor in chief was abruptly replaced. The fielding of the German Longitudinal Election Study Panel [@GLES2017Panel] in the same period provides the rare opportunity to precisely assess the effect of framing on political attitudes using a *difference-in-differences (DiD) design*.

The results indicate that even this strong increase in the use of crime frames did not affect consumers' immigration attitudes. The estimated effect is statistically indistinguishable from zero and significantly smaller than even minimally relevant effect magnitudes. Related issue attitudes, voting behavior, and issue importance are similarly unaffected, as is the association of immigration attitudes with the frame provided. Tracing different assumptions underlying the generalization of experimental estimates of framing effects into the real world implies that this lack in responsiveness is not an outcome of readers abandonment of the newspaper, nor the "crowding out" of information through other sources, but rather the crystallization of immigration attitudes prior to the change in framing.

Overall, these findings fundamentally challenge the external validity of a large body of framing effects research and suggest that the influence of media elites on public opinion is more limited than previously thought. Even a strong change in news framing did not affect political attitudes. Instead, it seems that the news media faces limited windows of opportunity to shape public opinion during the politicization of novel political issues. Additionally, the study contributes to a large body of media effects research in economics and political science by opening a research agenda to study the effects of political communication in the real world.

# Framing Effects on Public Opinion {#framing-effects}

@@

Work using the emphasis framing concept of framing mostly builds on the expectancy-value model developed by @Ajzen2000. This model suggests that an attitude on a given issue is a function of two things: the value of considerations relating to that issue and the strength of their association with the issue. The evaluation of the issue is subdivided into a number of considerations, which will carry clear evaluations. For example, to form an opinion about whether immigration to one's country should be restricted, a person might consider the humanitarian conditions in the countries of origin (negative consideration regarding restrictive migration policy), as well as the risk of increased crime (positive consideration; note that the model does not differentiate whether a consideration is grounded in objective reality). Individuals weigh each of these considerations [so-called "frames in thought", @Druckman2001] to form an overall opinion on the subject.

Figure \ref{fig:cogStor} visualizes this logic. An individual holds an attitude $y$ about a given issue (bottom square). The issue is associated with a number of different considerations (top circles), each of which carries an evaluation $x_i$, e.g. a potential surge in crime is considered a negative outcome. Based on the strength of the association $w_i$ of a given consideration with the issue, a given consideration's evaluation is more or less reflected in the overall attitude $y$, which is equal to the weighted sum of evaluations. For example, if a given individual associates immigration strongly with an increase in criminal activity and welfare fraud, but only very weakly with humanitarian considerations, cultural enrichment, liberty, or economic growth, they will form a rather negative attitude about the issue, as the former factor more prominently in their reasoning about migration.

@@

# Limits of Framing Experiments {#framing-limits}

@@

Second, even if consumers receive the content, it is not clear how much an individual relies on this source when forming an opinion on the subject.  Information from different sources might 'crowd out' the framing of a single outlet: by consuming content from many sources, individuals give less attention to a given source, thereby severely limiting the overall impact of that outlet's coverage. Experimental evidence suggests that, when an individual receives competing frames, these cancel each other out, although the precise impact is dependent on the relative timing of exposure, the quality of the frame, as well as individual characteristics [@Chong2007b; @Chong2010; @Chong2013].

Third, cognitive engagement with frames might strongly differ in the real world compared to experimental contexts. Once individuals formed political attitudes, they might resist framing: Motivated reasoners presented with a plethora of news content about an issue will spend less time and cognitive energy on counter-attitudinal arguments than those supporting their viewpoint, lending more importance to considerations raised by amenable content [@Taber2006]. Experimental evidence suggests similar dynamics specifically for framing: when individuals are free to choose content, initial frames dominate the opinion-formation process, while novel frames are rejected [@Druckman2012Bias] and even serve to strengthen pre-existing attitudes [@Bechtel2015]. This also explains why citizens are less reactive to framing attempts on salient and politicized issues [@Lecheler2009b].

@@

# Existing Observational Studies of Framing Effects

Despite repeated calls for the assessment of framing effects beyond experimental studies [@Busby2019; @Kinder2007; @Leeper2020], there is little research addressing this concern. While a large body of observational media effects research exists, it has not paid specific interest to framing, instead assessing the impact of changing news exposure on issue attitudes and voting behavior more broadly. Most of these studies are concerned with exposure to a different outlet [@Bail2018; @Broockman2022; @Foos2022; @Grossman2022; @Guess2021]. This is problematic, as citizens are often aware of outlets' political slant and engage with the coverage differently [@Baum2008a; see also @Spirig2020 for a discussion of this issue]. Those studies focusing on within-outlet changes in coverage do not identify changing framing directly and assess effects on a broad set of dependent variables, rather than specific issue attitudes [@Levendusky2022; @Spirig2020].

The precious small body of existing observational research specifically about framing is often riddled with endogeneity issues. For example, @Jerit2008 investigates the effects of news framing on public opinion. She finds that framing is less effective in affecting public opinion compared to engagement with opposing considerations, a finding that suggests that framing is rather ineffective. Others show with similar designs that only specific frames seem to drive issue importance [@Mclaren2018], but do not assess effects on attitudes. However, these studies restrict themselves to the assessment of correlational evidence. This is especially problematic regarding framing effects, as news content can affect consumers' inclination to consume news (see preceding section), and political actors might react to public opinion. As both variables affect each other at the same time, it is impossible to assess which direction causality runs.

To my knowledge, there are only two observational studies assessing the causal effect of exposure to specific content on political attitudes. @Slothuus2010b takes advantage of a sudden shift in a political parties' framing of retirement benefits, and shows that this changed framing has substantial impact on party supporters' attitudes about the issue. However, given that a positional change coincided with the changing framing of the issue and parties' special role in citizens' information process, one cannot distinguish the effect of partisan cue-taking from the persuasive effect of the frame itself. Additionally, it is unclear how these findings translate from parties' communication to news framing effects.

The only study which - to my knowledge - assesses the precise impact of changing news content on political attitudes comes from @Djourelova2020. She exploits a ban of the term "illegal migrant" in the *Associated Press* and shows that this language has modest effects on migration attitudes. While the study itself reveals impressively how language can shape political attitudes, it constitutes a highly specific case and a rather minor change in the framing of migration. To gain a deeper understanding of how specifically *framing* affects immigration attitudes, it is necessary to assess broader changes in the representation of the immigration issue.

Overall, it becomes obvious that we lacks an understanding of how news framing affects political attitudes in the real world. While experimental studies find strong effects, several assumptions have to be made to translate them beyond the lab or survey context. Observational evidence on media effects is only informative about broadly defined 'media effects' in specific cases, mostly without a more detailed focus on which content affects political attitudes. In order to identify the precise effects of changing news content, it is necessary to assess within-outlet changes in news coverage and their effect on specific issue attitudes using a design able to identify both the precise nature of the changes in content and their causal effects on attitudes. The following section therefore presents a research design building on a rare change in the migration framing in Germany's largest tabloid, *Bild*. It then develops a framework to measure emphasis framing in political communication, and presents a panel dataset tracking migration attitudes in the time-frame of interest, as well as a *difference-in-difference* design to estimate the causal effect of the changing framing. The final section discusses the expected effect size from experimental studies, in order to provide a baseline for comparison.

@@

In the summer of 2015, Germany became sanctuary for around 800,000 refugees who had fled war, violence, and famine in Syria and elsewhere. The German public debated how to deal with the newly arriving and for the first time since the post-war years, a radical right party was likely to win representation in the national parliament. Surprisingly, given its traditional populist style and right-wing takes on migration, the major tabloid newspaper *Bild* initially maintained a somewhat balanced coverage of refugees, which - although still portraying migrants more negatively than other outlets [@Maurer2019] - even started a campaign called "#RefugeesWelcome" to gather support for the newly arriving\footnote{See \url{https://web.archive.org/web/20160114154402/https://www.bild.de/news/topics/fluechtlingshilfe/wir-helfen-buehne-42385428.bild.html}.}. This was often connected to chief editor Kai Diekmann, who hosted a refugee family in his home [@BR2015Diekmann].

About one and a half years later, in December 2016, Axel Springer, the company owning *Bild*, announced that Diekmann will leave the company. Less than a week after the announcement, it became public that prosecutors investigate Diekmann for sexual harassment. While the company claimed that this was not the reason for Diekmann's departure, the process might have been sped up as a result of the investigations\footnote{According to a former editor of the newspaper I talked to, rumors in the company at the time support this interpretation of events.} [@Spiegel2017Diekmann]. The migration coverage in *Bild* severely changed when Diekmann left the newspaper in February 2017 and Julian Reichelt took over the position as editor-in-chief. The paper increasingly started painting a picture of refugees as criminals on the tabloid's front page [@Niggemeyer2018Bild; @DLF2018Reichelt]. Headlines like "I killed 40 people and want asylum" increasingly dominated the migration coverage of the newspaper. In a speech months before he took over the position as chief editor, Reichelt outlined his motivation for such a shift in migration coverage: "I can assure you: nothing has hurt [*Bild*] economically as much as our clear, humane, empathetic stance in the refugee crisis" \footnote{Reichelt in a speech at Deutschlandfunk's conference "Formate des Politischen 2016": \url{https://vimeo.com/190347766}}. While he expressed his support for the decision to take this stance at the time, he argued that the German media shows too little tolerance towards anti-immigration attitudes, thereby excluding those holding such views from political discourse.

This case presents a unique opportunity to study the effects of media content on public opinion. As I will show below, the emphasis of crime strongly increased in *Bild*'s migration coverage following Reichelt's takeover. The change of editor itself was not discussed in the media, which means any opinion changes among readers can only be the result of the changing content [see @Spirig2020]. Additionally, the timing of the editorial change is potentially exogenous, given the pressure from allegations against Diekmann. In summary, *this case is as close as possible to a field experiment in which a specific consideration of migration coverage is amplified in the coverage of one newspaper, but not others*.

In the following sections, I will outline a framework for the measurement of emphasis frames in communication and showcase the increasing emphasis of crime in *Bild*'s migration coverage (treatment). Subsequent sections will describe the panel data and variables to measure frames in thought (namely migration attitudes and their correlation with attitudes about crime) and define the difference-in-difference estimator. I will close with a discussion of the expected experimental effect size and the minimal effect size of interest to provide baselines for comparison of the estimator.

@@

To measure migration framing in the news, I collected 2.5 million newspaper articles from the websites of the most important German broadsheets *Frankfurter Allgemeine Zeitung (FAZ), Spiegel Online (SPON), Süddeutsche Zeitung (SZ), Die Tageszeitung (TAZ), Die Welt*, as well as *Bild* for the period 2013-2019. These outlets reach a substantial share of the German public, with *Bild* distributing over 1.7 million print copies and registering over 550 million web impressions at the time (see section 1.1 in the appendix for more information\footnote{Online appendix available at \url{https://github.com/nicolaiberk/bild/raw/main/paper/appendix.pdf}.}). Following the value-expectancy model, I identify crime frames in migration coverage in a two-step process: first, I use supervised machine learning to identify content about migration. Second, I identify content about crime *within* migration coverage by training and applying another supervised classifier. This approach closely resembles the value-expectancy model outlined in section \@ref(framing-effects), as I first identify coverage about the *issue*, before identifying the consideration emphasized in the discussion of this issue.

To identify whether an article is about migration, I employ supervised machine learning. As migration content is rare, and sufficient training data needs to be available to train functional classifiers, articles cannot be selected randomly for annotation\footnote{If migration content made up about 3 percent of all articles and the sample were selected randomly, one would find around 30 migration articles per 1000 articles, which provides rather limited information about the class of interest and results in worse recall of the classifier, as the best guess in highly imbalanced data is simply the most common outcome.}. Instead, I take a more informed approach by oversampling articles which are likely about migration: after constructing a seed dictionary of migration-related terms, I apply dictionary extension using German GloVe word-embeddings\footnote{Downloadable from \url{https://deepset.ai/german-word-embeddings}.} to construct a comprehensive dictionary, and apply it to the articles\footnote{The dictionary can be found in appendix XX.}. Based on the relative share of migration words in an article, I draw a stratified sample of 300 articles (100 from articles with no migration terms, 100 from the quarter with the highest share of migration terms and 100 from the remaining articles with relative frequencies of migration-related terms in between these two groups) for each newspaper for a total of 1,800 papers. Then, a coder hand-coded these articles, assessing whether their main topic is related to migration\footnote{The coder was instructed to annotate articles referring to migration or integration, which includes discussions of refugee numbers, immigrants' and naturalization rights, and general reports of migration.}.

Using this sample, a BERT transformer model\footnote{\url{https://huggingface.co/bert-base-german-cased}} is fine-tuned on a random subset of 1,400 annotated articles (training set) with a validation set of 200 articles. After fine-tuning, the model correctly classifies 95.5% of a test set of 200 articles (F1: 0.94, recall: 0.93, precision: 0.95). This classifier is then used to annotate all 2.5 million newspaper articles. For the period from 2013 to 2019, around 90,000 articles (3.6%) are identified to treat the issue of migration\footnote{A precise description of the classification process for migration and crime-related articles, including sampling dictionaries, instructions for coders, and preprocessing steps, can be found in appendix 1.2.}.

To identify migration coverage containing the crime consideration and thereby associating migration with crime, I first fine tune a BERT classifier for German language in a similar fashion as above, this time with coverage about crime as the outcome of interest. Using a training set of 1400 articles (sleected using the identlical process as above), a validation set of 200 and a test set of 200 articles, the classifier reaches a very good performance with an F1 score of 85.7%. I then apply this classifier to the approximately 90,000 articles about migration identified by the first classifier. About 7% of all migration coverage consists of articles about some form of crime. Exemplifying the nature of the treatment, table \@ref(tab:exampleArticles) displays the headlines of the five migration articles within *Bild* which were identified to discuss crime with the highest likelihood. All articles refer to crimes of migrant groups, with two referring to violence in refugee homes, two discussing police raids regarding asylum fraud, and one discussing a police report about the sexual assaults surrounding new years eve in cologne.

The left panel of figure \@ref(fig:treatment) shows the share of migration coverage devoted to crime news (solid, black line), once for Bild and once for all other newspapers (dashed, gray line). As we can see, *Bild*'s attention to crime news in its migration coverage substantially increased directly following the takeover of Julian Reichelt, while it focused less on the issue preceding the editorial change\footnote{Note that this data consists exclusively of original Bild content. Treatment estimates including third-party authored content are similar in size. For a more extensive discussion see appendix XX.}. The trends run parallel to each other preceding the treatment. The lower attention to crime in 2015 and the increase in 2016 is in line with prior research [@Maurer2019]. The right-hand panel shows difference-in-difference estimates of the crime emphasis in migration coverage of one newspaper compared to all others, before and after the editorial change in *Bild*. One can see that *Bild* was the only newspaper which substantially changed its emphasis of the frame, increasing the share of articles devoted to crime by around 5 percentage points, while other newspapers maintained their levels of attention.

>>>>

---
title: "Limits of Media Effects"
subtitle: "How Politicisation Shapes the Pace of Attitudinal Change"
author: "Nicolai Berk"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    fig_caption: yes
    number_sections: false
    includes:
      in_header: preamble.tex
# subtitle: Manuscript prepared for the 12th Annual Conference of the European Political Science Association in Prague, Czech Republic
fontsize: 12pt
geometry: margin=1.5in
abstract: "What are the limits of the news media's ability to shape public opinion? I argue that politicization conditions the responsiveness of individual attitudes to changes in the information environment. Exploiting a rare change in the migration framing of the biggest German tabloid *Bild*, I estimate the precise impact of this shift on immigration attitudes using large-scale panel data. Despite a 42 percent increase of the emphasis of crime in migration coverage, I find a robust null effect on immigration attitudes and a number of related variables. Further analyses suggest that media outlets' ability to manipulate public opinion in the short term seems to be limited to brief opportunity windows during the politicisation of a given issue."
keywords: "Public Opinion, Media Effects, News Framing, Opinion Formation, Emphasis Framing, Conditions of Media Effects."
bibliography: C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib
nocite: |
  @CologneReport2016
indent: true
---

```{r setup, include = F, cache=F}

knitr::opts_chunk$set(echo = F, message = F, warning = F, cache = F, fig.height=5, fig.width=8, fig.align='center', eval = T, fig.pos = "!ht", out.extra = "")

library(tidyverse)
library(dplyr)
library(data.table)
library(fixest)
library(lubridate)
library(patchwork)
library(here)
library(extrafont)

loadfonts(device = "win")

@@

\doublespacing

# Introduction

A long tradition of research on public opinion suggests that political attitudes are a direct reflection of the information environment shaping the 'pictures in our heads' [@Lippmann1922]. This view is supported by an extensive body of experimental evidence showing that framing is highly effective in the manipulation of issue attitudes. However, prior work has shown that the extrapolation of experimental results to real-world environments is often difficult [@Barabas2010], and that the effects of framing are highly dependent on the political context [@Chong2013]. Media effects research has established that real-world changes in the media environment *can* affect issue attitudes [@Foos2022; @Grossman2022; @Djourelova2020], but not always [@Chiang2011a; @Spirig2020]. We currently know little about the real-world conditions under which news framing affects public opinion.

This study attempts to address this research gap presenting evidence from Germany, 2017-2020. This period - following a spike in asylum applications in 2015 - was marked by a focus on immigration in political discourse, as well as the first parliamentary entrance of a radical-right party since the post-war years. Many observers connected this sudden change in Germany's political landscape to a changing immigration discourse and the normalization of far-right ideas in public debate [@NYT2017BTW]. Germany's largest tabloid newspaper, *Bild*, was seen as a strong driver of this development: it was claimed that by framing asylum seekers as criminals, the paper fueled anti-immigration sentiment and support for the radical-right *AfD* [@Niggemeyer2018Bild].

In the remainder of this paper, I will argue that *Bild* missed a critical window of opportunity to shape public opinion. As readers' immigration attitudes had crystallized following the politicization of immigration in 2015, their immigration attitudes remained resistant to a change in the tabloid's migration framing. I support my argument with evidence presented in three steps: first, I show that *Bild* indeed increasingly emphasized crime-related news in its migration coverage, following an exogenously timed change of the editor-in-chief. I then exploit this natural experiment to provide evidence that this change did little to affect readers' immigration attitudes or related variables. In a last step, I trace different explanations for the absence of this effect, suggesting that the pace of attitudinal change severely decreased following the politicization of the issue in 2015.

To assess whether the migration framing in *Bild* really changed, I develop an approach to measure framing in political texts. I apply this framework to a dataset of 2.5 million articles from the major German daily newspapers. Using supervised classifiers, I first select migration-related coverage, to then estimate the prevalence of crime content within this coverage. This analysis suggests that following the change of the editor-in-chief, the emphasis of crime in migration content increased by over 40 percent.

To estimate the effect of this changing framing on immigration attitudes, I employ panel data from the German Longitudinal Election Study, fielded in the period of interest [@GLES2017Panel]. I use a *difference-in-differences* (DiD) design to estimate the impact of the increased emphasis of crime in the coverage of migration and refugees. Results show that the changing framing did not result in more restrictive attitudes towards immigration, nor did it affect the association of crime with migration, views on integration, or the perceived importance of the immigration issue.

Lastly, I explore different explanations why this shift in migration framing did not bring about attitudinal changes. First, I assess whether readers with initially liberal migration attitudes selected out of the newspaper in response to the changing portrayal of migrants. Reported readership of respondents with liberal migration attitudes remains at similar levels as that of readers with restrictive migration attitudes. Second, I investigate whether readers with many different sources of information are less reactive to the changing coverage and mask the effect. I find no effect heterogeneity across the reported consumption of newspaper, TV and online news, ruling out the possibility that readers with more diverse information hide an effect among exclusive *Bild* readers. Third, I test whether attitudes change after longer-term exposure. I find small effects after more than one and a half years, suggesting a very slow pace of attiudinal change. Lastly, I assess whether attitudes crystallized after the politicization of immigration in 2015. Using data from a long-term online panel [@GLES2018LangfristPanel], I show that the average individual change in immigration attitudes was substantially lower in the period 2016-2018 compared to the period 2014-2016.

This evidence suggests a general model of opinion change that contributes to distinct literatures on media effects, attitude formation, and framing. Students of the effects of news content on political attitudes have started to explore the scope conditions of media effects [@Chiang2011a; @Spirig2020]. This paper contributes to this literature by showing that changes in media coverage are ineffective when they are not addressing currently politicised issues. As a result, elite influence on public opinion is limited outside specific windows of opportunity.

The analyses in this paper also speak to recent evidence of overall attitude stability [@Kustov2021]: citizens' immigration attitudes *are* rather stable most of the time. However, my findings suggest that the pace of attitudinal change increases dramatically when an issue becomes salient in public debate. The emerging picture is an ebb and flow of attitudinal stability and change, dependent on the salience of the issue at stake.

Lastly, the evidence speaks to the study of framing effects. Fifteen years ago, Donald Kinder called for observational research on framing effects: "Experimental results can always be questioned on their generalizability, and framing effects are no exception. [...] A more balanced reading of frame effects requires methodological diversification, experiments and studies oriented to the world outside" [-@Kinder2007, 157]. By precisely identifying the nature of the change in news coverage, this study answered this call. The null effects found suggest that framing effects do not easily generalize from the experimental context.

The paper proceeds as follows: I provide a brief review of the expectancy-value model of attitude formation, which informs my hypotheses and content analysis, before discussing potential limits of framing effects in real-world environments. I then present the research design, content analysis, and a discussion of expected and minimal effect size of interest, before turning to the estimation of the effect of changing immigration framing on immigration attitudes and trace different explanations for the observed resistance. In the conclusion, I outline the implications of this model for framing, media, and public opinion research, as well as the limitations of this study.

# The Expectancy-Value Model {#framing-effects}

@@

Work using the emphasis framing concept of framing mostly builds on the expectancy-value model developed by @Ajzen2000. This model suggests that an attitude on a given issue is a function of two factors: the value of considerations relating to that issue and the strength of their association with the issue. The evaluation of the issue is subdivided into a number of considerations, which will carry clear evaluations. For example, to form an opinion about whether immigration to one's country should be restricted, a person might consider the humanitarian conditions in the countries of origin (negative consideration regarding restrictive migration policy), as well as the risk of increased crime (positive consideration; note that the model does not differentiate whether a consideration is grounded in objective reality). Individuals weigh each of these considerations [so-called "frames in thought", @Druckman2001] to form an overall opinion on the subject.

Figure \ref{fig:cogStor} visualizes this logic. An individual holds an attitude $y$ about a given issue (bottom square). The issue is associated with a number of different considerations (top circles), each of which carries an evaluation $x_i$, e.g. a potential surge in crime is considered a negative outcome. Based on the strength of the association $w_i$ of a given consideration with the issue, that consideration's evaluation is more or less reflected in the overall attitude $y$, which is equal to the weighted sum of evaluations. For example, if a given individual associates immigration strongly with an increase in criminal activity and welfare fraud, but only very weakly with humanitarian considerations, cultural enrichment, liberty, or economic growth, they will form a rather negative attitude about the issue, as the former factor more prominently in their reasoning about migration.

@@

# Limits of Framing Effects {#framing-limits}

@@

Second, even if consumers receive the content, it is not clear how much an individual relies on this source when forming an opinion on the subject.  Information from different sources might 'crowd out' the framing of one specific outlet: by consuming content from many sources, individuals give less attention to a given source, thereby severely limiting the overall impact of that outlet's coverage. Experimental evidence suggests that, when an individual receives competing frames, these cancel each other out, although the precise impact is dependent on the relative timing of exposure, the quality of the frame, as well as individual characteristics [@Chong2007b; @Chong2010; @Chong2013]. Additionally, individuals might update their perceptions slowly, as the increased emphasis of a consideration is not necessarily evident from the first few articles, instead arising from repeated exposure over time. Especially given that this study is concerned with a change in the *relative* emphasis of a consideration, it might take some time for citizens' perceptions to catch up to the changing issue representation.

Third, cognitive engagement with frames might strongly differ in the real world compared to experimental contexts. Once individuals formed political attitudes, they might resist framing: Citizens presented with a plethora of news content about an issue will spend less time and cognitive energy on counter-attitudinal arguments than those supporting their viewpoint, lending more importance to considerations raised by amenable content [@Taber2006]. Experimental evidence suggests similar dynamics specifically for framing: when individuals are free to choose content, initial frames dominate the opinion-formation process, while novel frames are rejected [@Druckman2012Bias] and even serve to strengthen pre-existing attitudes [@Bechtel2015].

@@

@@

In the summer of 2015, Germany became sanctuary for around 800,000 refugees who had fled war, violence, and famine in Syria and elsewhere. The German public debated how to deal with the newly arriving and for the first time since the post-war years, a radical right party was likely to win representation in the national parliament. Surprisingly, given its traditional populist style and right-wing takes on migration, the major tabloid newspaper *Bild* initially maintained a somewhat balanced coverage of refugees, which – although still portraying migrants more negatively than other outlets [@Maurer2019] – even started a campaign called "#RefugeesWelcome" to gather support for the newly arriving\footnote{See \url{https://web.archive.org/web/20160114154402/https://www.bild.de/news/topics/fluechtlingshilfe/wir-helfen-buehne-42385428.bild.html}.}. This was often connected to chief editor Kai Diekmann, who hosted a refugee family in his home [@BR2015Diekmann].

About one and a half years later, in December 2016, Axel Springer, the company owning *Bild*, announced that Diekmann would leave the company. Less than a week after the announcement, it became public that prosecutors investigated Diekmann for sexual harassment. While the company claimed that this was not the reason for Diekmann's departure, the process might have been sped up as a result of the investigations\footnote{According to a former editor of the newspaper I talked to, rumors in the company at the time support this interpretation of events.} [@Spiegel2017Diekmann]. The migration coverage in *Bild* severely changed when Diekmann left the newspaper in February 2017 and Julian Reichelt took over the position as editor-in-chief. The paper increasingly started painting a picture of refugees as criminals on the tabloid's front page [@Niggemeyer2018Bild; @DLF2018Reichelt]. Headlines like "I killed 40 people and want asylum" increasingly dominated the migration coverage of the newspaper. In a speech months before he took over the position as chief editor, Reichelt outlined his motivation for such a shift in migration coverage: "I can assure you: nothing has hurt [*Bild*] economically as much as our clear, humane, empathetic stance in the refugee crisis" \footnote{Reichelt in a speech at Deutschlandfunk's conference "Formate des Politischen 2016": \url{https://vimeo.com/190347766}}. While he expressed his support for the decision to take this stance at the time, he argued that the German media shows too little tolerance towards anti-immigration attitudes, thereby excluding those holding such views from political discourse.

This case presents a unique opportunity to study the effects of media content on public opinion. As I will show below, the emphasis of crime strongly increased in *Bild*'s migration coverage following Reichelt's takeover. The change of editor itself was not discussed in the media, which means any opinion changes among readers can only be the result of the changing content [see @Spirig2020]. Additionally, the timing of the editorial change is likely exogenous, given the pressure from allegations against Diekmann. In summary, *this case is as close as possible to a field experiment in which a specific consideration of migration coverage is amplified in the coverage of one newspaper, but not others*.

In the following sections, I will outline a framework for the measurement of emphasis frames in communication and showcase the increasing emphasis of crime in *Bild*'s migration coverage. Subsequent sections will describe the panel data and variables to measure frames in thought (namely migration attitudes and their association with attitudes about crime) and define the difference-in-difference estimator. I will close with a discussion of the expected experimental effect size and the minimal effect size of interest to provide baselines for comparison of the estimator.

@@

To measure migration framing in the news, I collected 2.5 million newspaper articles from the websites of the most important German broadsheets *Frankfurter Allgemeine Zeitung (FAZ), Spiegel Online (SPON), Süddeutsche Zeitung (SZ), Die Tageszeitung (TAZ), Die Welt*, as well as *Bild* for the period 2013-2019. These outlets reach a substantial share of the German public, with *Bild* distributing over 1.7 million print copies and registering over 550 million web impressions at the time (see section 2.1 in the appendix for more information\footnote{Online appendix available at \url{https://github.com/nicolaiberk/bild/raw/main/paper/appendix.pdf}.}). Following the value-expectancy model, I identify crime frames in migration coverage in a two-step process: first, I use supervised machine learning to identify content about migration. Second, I identify content about crime *within* migration coverage by training and applying another supervised classifier. This approach closely resembles the value-expectancy model outlined [above](#framing-effects), as I first identify coverage about the *issue*, before identifying the consideration emphasized in the discussion of this issue.

To identify whether an article is about migration, I employ supervised machine learning. As migration content is rare, and sufficient training data needs to be available to train functional classifiers, articles cannot be selected randomly for annotation\footnote{If migration content made up about 3 percent of all articles and the sample were selected randomly, one would find around 30 migration articles per 1000 articles, which provides rather limited information about the class of interest and results in worse recall of the classifier, as the best guess in highly imbalanced data is simply the most common outcome.}. Instead, I take a more informed approach by oversampling articles which are likely about migration: after constructing a seed dictionary of migration-related terms, I apply dictionary extension using German GloVe word-embeddings\footnote{Downloadable from \url{https://deepset.ai/german-word-embeddings}.} to construct a comprehensive dictionary, and apply it to the articles\footnote{The dictionary can be found in appendix 2.2.1.}. Based on the relative share of migration words in an article, I draw a stratified sample of 300 articles (100 from articles with no migration terms, 100 from the quarter with the highest share of migration terms and 100 from the remaining articles with relative frequencies of migration-related terms in between these two groups) for each newspaper for a total of 1,800 papers. Then, a coder hand-coded these articles, assessing whether their main topic is related to migration\footnote{The coder was instructed to annotate articles referring to migration or integration, which includes discussions of refugee numbers, immigrants' and naturalization rights, and general reports of migration. More information can be found in Appendix 2.2.3.}.

Using this sample, a BERT transformer model\footnote{\url{https://huggingface.co/bert-base-german-cased}}, pre-trained on a large German corpus of Wikipedia articles, News, and legal documents, is fine-tuned on a random subset of 1,400 annotated articles (training set) with a validation set of 200 articles. After fine-tuning, the model correctly classifies 95.5% of a test set of 200 articles (F1: 0.94, recall: 0.93, precision: 0.95). This classifier is then used to annotate all 2.5 million newspaper articles. For the period from 2013 to 2019, around 90,000 articles (3.6%) are identified to treat the issue of migration\footnote{A precise description of the classification process for migration and crime-related articles, including sampling dictionaries, instructions for coders, and preprocessing steps, can be found in appendix 2.2.}.

To identify migration coverage containing the crime consideration and thereby associating migration with crime, I first fine tune a BERT classifier for German language in a similar fashion as above, this time with coverage about crime as the outcome of interest. Using a training set of 1400 articles (selected using the identical process as above), a validation set of 200 and a test set of 200 articles, the classifier reaches a very good performance with an F1 score of 85.7%. I then apply this classifier to the approximately 90,000 articles about migration identified by the first classifier. About 7% of all migration coverage consists of articles about some form of crime\footnote{See appendix 2.2.7 for more information.}.

The left panel of figure \@ref(fig:treatment) shows the share of migration coverage devoted to crime news, once for *Bild* (solid, black line) and once for all other newspapers (dashed, gray line). As we can see, *Bild*'s attention to crime news in its migration coverage substantially increased directly following the takeover of Julian Reichelt, while it focused less on the issue preceding the editorial change\footnote{Note that this data consists exclusively of original Bild content. Treatment estimates including third-party authored content are similar in size. For a more extensive discussion see appendix 2.2.5.}. The trends run parallel to each other preceding the treatment. The lower attention to crime in 2015 and the increase in 2016 is in line with prior research [@Maurer2019]. The right-hand panel shows difference-in-difference estimates of the crime emphasis in migration coverage of one newspaper compared to all others, before and after the editorial change in *Bild*. One can see that *Bild* was the only newspaper which substantially changed its emphasis of the frame, increasing the share of articles devoted to crime by around 5 percentage points (over 40 percent percent of the pre-treatment average), while other newspapers maintained their levels of attention.
